Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2023-07-05 popl24/scratch.tex
checkpoint

\begin{lemma}
  \label{lemma-separation}
  The following properties hold:
  \begin{enumerate}
  \item $\sep{\pmf}{Y}{Z}$ iff $\sep{\pmf}{Z}{Y}$
  \item $\vc{\pmf}{x}{y}$ if  $\vc{\pmf}{y}{x}$
  \item $\vc{\pmf}{x}{y}$ and $\vc{\pmf}{y}{z}$ imply $\vc{\pmf}{x}{z}$
  \item ($\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z}$) iff $\sep{\pmf}{X}{(Y \cup Z)}$
  %\item If $\prog_1;\prog_2$ is safe and $\vars(\prog_1) \cap \vars(\prog_2) = \varnothing$
  %  then $\sep{\progtt(\prog_1;\prog_2)}{\vars(\prog_1)}{\vars(\prog_2)}$.
  \end{enumerate}
\end{lemma}

@@

\begin{definition}[GMW Gate Certification]
  \label{definition-gmwgate-certification}
  Let $\mathit{gate}$ be a gate function and let $e$ be the following
  $\metaprot$ program:
  $$
  \begin{array}{l}
    \ttt{let in1 = \{c1 = flip[1,s11];c2 = flip[2,s12]\} in}\\
    \ttt{let in2 = \{c1 = flip[1,s21];c2 = flip[2,s22]\} in}\\
    \mathit{gate}\ttt{(g,in1,in2)}
  \end{array}
  $$
  and let $
  \config{\varnothing}{e}\redxs\config{\prog}{\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}}
  $.
  Then $\mathit{gate}$ is \emph{certified} iff the following conditions hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21]\}}}{\ttt{\{v[2,gout]\}}}$
  \item $\sep{\progtt(\prog)}{\ttt{\{flip[2,s12],flip[2,s22]\}}}{\ttt{\{v[1,gout]\}}}$
  \item All variables in:$$\vars(\prog) - \{ \ttt{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]}\}$$
    are distinguished by $\ttt{g}$ (i.e., contain $\ttt{g}$ as an identifier substring). 
  \end{enumerate}
\end{definition}

\begin{lemma}
  \label{lemma-gmw-frame}
  If $\sep{\progtt(\prog_2)}{Y}{Z}$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
  $X \not\in \vars(\prog_2)$ then $\sep{\progtt(\prog_1;\prog_2)}{X}{Z}$. 
\end{lemma}

@@

\begin{lemma}[GMW Gate Invariant]
  \label{lemma-preservation-gmw}
  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ and certified $\mathit{gate}$ where
  for some $\ttt{g1}$ and $\ttt{g2}$, :
  $$
  \begin{array}{lcr}
   v_1 &=& \ttt{\{c1 = v[1,g1out]; c2 = v[2,g1out]\}}\\
   v_2 &=& \ttt{\{c1 = v[1,g2out]; c2 = v[2,g2out]\}}
  \end{array}
  $$
  and let:
  $$
  \config{\prog_1}{E[\mathit{gate}(\ttt{g},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}]}
  $$
  If we have the following precondition, where $\iov(\prog_1) = S \cup V$:
  $$
  \sep{\progd(\prog_1)}{(S_{\{2\}} \cup V_{\{2\}})}{(S_{\{1\}} \cup V_{\{1\}})}
  $$
  then we have as a postcondition:
  $$
  \sep{\progd(\prog_1)}{(S_{\{2\}} \cup V_{\{2\}} \cup \{ \ttt{v[2,gout]} \})}{(S_{\{1\}} \cup V_{\{1\}} \cup \{ \ttt{v[1,gout]} \})}
  $$
\end{lemma}

\begin{proof}
  Let $\prog$ be as defined in Definition \ref{definition-gmwgate-certification}.
  We observe:
  $$
  {\footnotesize
    \begin{array}{c}
      \prog_2 = \\
      \prog[\ttt{v[1,g1out]}/\ttt{flip[1,s11]}][\ttt{v[2,g1out]}/\ttt{flip[2,s12]}][\ttt{v[1,g2out]}/\ttt{flip[1,s21]}][\ttt{v[2,g2out]}/\ttt{flip[2,s22]}]
    \end{array}
  }
  $$ Now, the assumption of well-typedness of
  $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ assures that $\ttt{g}$
  has not been previously used as a gate identifier in $\prog_1$,
  since otherwise the views $\ttt{v[1,gout]}$ and $\ttt{v[2,gout]}$
  would have been previously defined. This ensures $\vars(\prog_1,x)
  \cap \vars(\prog_2) = \varnothing$ for:
  $$x \in \{ \ttt{v[1,g1out]}, \ttt{v[2,g1out]}, \ttt{v[1,g2out]}, \ttt{v[2,g2out]}\}$$
  given condition (iii) in Definition \ref{definition-gmwgate-certification}.
  Thus by Lemma \ref{lemma-substitution} we have:
  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{\ttt{\{v[2,gout]\}}}$$
  and
  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout]\}}}$$
  Since $\vars(\prog_1) \cap \vars(\prog_2)$ are only the wire inputs to $\ttt{g}$,
  therefore by assumption and Lemma \ref{lemma-gmw-frame}:
  $$\sep{\progtt(\prog_1;\prog_2)}{(S_{\{1\}} \cup V_{\{1\}})}{\ttt{\{v[2,gout]\}}}$$
  and also:
  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{(S_{\{2\}}\cup V_{\{2\}})}$$
  etc., so the result follows. 
\end{proof}

\begin{definition}[GMW Encode Certification]
  Let $\mathit{encode}$ be an encoding function, and let:
  $$
  \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
  \config{\prog}{\ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}}
  $$
  where:
  $$
  \begin{array}{rcl}
    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
  \end{array}
  $$
  Then $\mathit{encode}$ is \emph{certified} iff each of the following conditions hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\sep{\progtt(\prog)}{\ttt{\{ s[1,s1] \}}}{\ttt{\{ v[2,s1out], v[2,s2out] \}}}$
  \item $\sep{\progtt(\prog)}{\ttt{\{ s[2,s2] \}}}{\ttt{\{ v[1,s1out], v[1,s2out] \}}}$
  \item All $x \in \vars(\prog)$ are distinguished by $\ttt{in1}$ or $\ttt{in2}$. 
  \end{enumerate}
\end{definition}

\begin{theorem}
  \label{theorem-gmw-NIMO}
  If $e$ is a well-typed GMW circuit definition using certified
  components and $\config{\varnothing}{e} \redxs
  \config{\prog}{\varnothing}$, then $\NIMO(\prog)$.
\end{theorem}
\begin{proof}
  Let $\iov(\prog) = S \cup (V \cup \outv)$. Noting that any encoding
  necessarily uses entirely fresh variables by Definition and type
  safety, by applying Lemmas \ref{lemma-compose-gmw} and
  \ref{lemma-wire-framing} and the frame rule of
  \cite{barthe2019probabilistic} we obtain:
  $$
  \sep{\progd(\prog)}{(S_{\{2\}}\cup\{\outv\})}{V_{\{1\}}}
  \quad \text{and} \quad \sep{\progd(\prog)}{(S_{\{1\}}\cup\{\outv\})}{V_{\{2\}}}
  $$
  That is, the marginal distribution of client 1's secrets and the output are independent
  of client 2's views, and vice-versa. The result follows. 
\end{proof}

>>>>

\begin{lemma}
  \label{lemma-separation}
  The following properties hold:
  \begin{enumerate}
  \item $\sep{\pmf}{Y}{Z}$ iff $\sep{\pmf}{Z}{Y}$
  \item $\vc{\pmf}{x}{y}$ if  $\vc{\pmf}{y}{x}$
  \item $\vc{\pmf}{x}{y}$ and $\vc{\pmf}{y}{z}$ imply $\vc{\pmf}{x}{z}$
  \item $\sep{\pmf}{X}{(Y \cup Z)}$ implies $(\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z})$
  \item $(\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z}$ and $\sep{\pmf}{Y}{Z})$ implies $\sep{\pmf}{X}{(Y \cup Z)}$ 
  %\item If $\prog_1;\prog_2$ is safe and $\vars(\prog_1) \cap \vars(\prog_2) = \varnothing$
  %  then $\sep{\progtt(\prog_1;\prog_2)}{\vars(\prog_1)}{\vars(\prog_2)}$.
  \end{enumerate}
\end{lemma}

@@

\begin{definition}[GMW Gate Certification]
  \label{definition-gmwgate-certification}
  Let $\mathit{gate}$ be a gate function and let $e$ be the following
  $\metaprot$ program:
  $$
  \begin{array}{l}
    \ttt{let in1 = \{c1 = flip[1,s11];c2 = flip[2,s12]\} in}\\
    \ttt{let in2 = \{c1 = flip[1,s21];c2 = flip[2,s22]\} in}\\
    \mathit{gate}\ttt{(g,in1,in2)}
  \end{array}
  $$
  and let $
  \config{\varnothing}{e}\redxs\config{\prog}{\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}}
  $.
  Then $\mathit{gate}$ is \emph{certified} iff the following conditions hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\sep{\progtt(\prog)}{\ttt{\{flip[1,s11],flip[1,s21]\}}}{\ttt{\{v[1,gout],v[2,gout]\}}}$
  \item $\sep{\progtt(\prog)}{\ttt{\{flip[2,s12],flip[2,s22]\}}}{\ttt{\{v[1,gout],v[2,gout]\}}}$
  % \item $\sep{\progtt(\prog)}{\ttt{\{v[1,gout]\}}}{\ttt{\{v[2,gout]\}}}$
  \item All variables in:$$\vars(\prog) - \{ \ttt{flip[1,s11],flip[1,s21],flip[2,s12],flip[2,s22]}\}$$
    are distinguished by $\ttt{g}$ (i.e., contain $\ttt{g}$ as an identifier substring). 
  \end{enumerate}
\end{definition}

\begin{lemma}
  \label{lemma-gmw-frame}
  If $\sep{\progtt(\prog_1;\prog_2)}{Y}{V}$ for $V \in \vdefs(\prog_2)$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
  $X \not\in \vars(\prog_2)$ then $\sep{\progtt(\prog_1;\prog_2)}{X}{V}$. 
\end{lemma}
\begin{proof}
  By assumption and Lemma \ref{lemma-sepjoin}, for all
  $\store^1,\store^2 \in \runs(\prog_1;\prog_2)$ we have $\store^1_Y
  \cap \store^2_V \subseteq \store$ for some $\store_a \in
  \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_Y \subseteq
  \store$ for some $\store_b \in \runs(\prog_1;\prog_2)$.  Thus, since
  we assume $X \not\in \vars(\prog_2)$ we have $\store_a \cap \store_b
  \in \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_V =
  \store_a \cap \store_b$, obtaining the result by
  \ref{lemma-sepjoin}.
\end{proof}

@@

\begin{lemma}[GMW Gate Invariant]
  \label{lemma-gmw-preservation}
  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ for certified $\mathit{gate}$ where:
  $$
  \config{\prog_1}{E[\mathit{gate}(\ttt{g},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
  $$
  If we have the following preconditions, where $\iov(\prog_1) = S \cup V$:
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$
  \item $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$
  \end{enumerate}
  then we have as a postconditions, where $\iov(\prog_1;\prog_2) = S \cup V'$:
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\sep{\progd(\prog_1;\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$
  \item $\sep{\progd(\prog_1;\prog_2)}{S_{\{1\}}}{V'_{\{2\}}}$
  \end{enumerate}
\end{lemma}

\begin{proof}
  Given well-typedness of $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$, we have for some $\ttt{g1}$ and $\ttt{g2}$:
  $$
  \begin{array}{lcl}
   v_1 &=& \ttt{\{c1 = v[1,g1out]; c2 = v[2,g1out]\}}\\
   v_2 &=& \ttt{\{c1 = v[1,g2out]; c2 = v[2,g2out]\}}\\
    v &=& \ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}
  \end{array}
  $$
  Let $\prog$ be as defined in Definition \ref{definition-gmwgate-certification}.
  We observe:
  $$
  {\footnotesize
    \begin{array}{c}
      \prog_2 = \\
      \prog[\ttt{v[1,g1out]}/\ttt{flip[1,s11]}][\ttt{v[2,g1out]}/\ttt{flip[2,s12]}][\ttt{v[1,g2out]}/\ttt{flip[1,s21]}][\ttt{v[2,g2out]}/\ttt{flip[2,s22]}]
    \end{array}
  }
  $$
  Now, the assumption of well-typedness also assures that $\ttt{g}$
  has not been previously used as a gate identifier in $\prog_1$,
  since otherwise the views $\ttt{v[1,gout]}$ and $\ttt{v[2,gout]}$
  would have been previously defined. This ensures $\vars(\prog_1,x)
  \cap \vars(\prog) = \varnothing$ for:
  $$x \in \{ \ttt{v[1,g1out]}, \ttt{v[2,g1out]}, \ttt{v[1,g2out]}, \ttt{v[2,g2out]}\}$$
  given condition (iii) of Definition \ref{definition-gmwgate-certification}.
  Thus by Lemma \ref{lemma-substitution} we have:
  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{\ttt{\{v[1,gout],v[2,gout]\}}}$$
  and
  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout],v[2,gout]\}}}$$
  And by Lemma \ref{lemma-gmw-frame} and condition (iii) of Definition
  \ref{definition-gmwgate-certification} we then have:
  $$\sep{\progtt(\prog_1;\prog_2)}{S_{\{1\}}}{\ttt{\{v[2,gout]\}}} \quad \text{and} \quad
  \sep{\progtt(\prog_1;\prog_2)}{S_{\{2\}}}{\ttt{\{v[1,gout]\}}}$$
  so by precondition assumptions and Lemma \ref{lemma-separation} we have:
  $$\sep{\progtt(\prog_1;\prog_2)}{S_{\{1\}}}{V'_{\{2\}}} \quad \text{and} \quad
  \sep{\progtt(\prog_1;\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$$
\end{proof}

\begin{definition}[GMW Encode Certification]
  \label{definition-gmwencode-certification}
  Let $\mathit{encode}$ be an encoding function, and let:
  $$
  \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
  \config{\prog}{\ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}}
  $$
  where:
  $$
  \begin{array}{rcl}
    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
  \end{array}
  $$
  Then $\mathit{encode}$ is \emph{certified} iff each of the following conditions hold:
  \begin{enumerate}[\hspace{5mm}i.]
    %\item $\sep{\progtt(\prog)}{\ttt{\{ s[1,s1],  v[1,s1out], v[1,s2out] \}}}{\ttt{\{ s[2,s2], v[2,s1out], v[2,s2out] \}}}$
  \item $\sep{\progtt(\prog)}{\ttt{\{s[1,s1]\}}}{\ttt{\{v[2,s1out],v[2,s2out]\}}}$
  \item $\sep{\progtt(\prog)}{\ttt{\{s[2,s2]\}}}{\ttt{\{v[1,s1out],v[1,s2out]\}}}$
  \item All $x \in \vars(\prog)$ are distinguished by $\ttt{in1}$ or $\ttt{in2}$. 
  \end{enumerate}

\end{definition}

\begin{lemma}[GMW Encode Invariant]
  \label{lemma-gmw-encode}
  Given well-typed $\prog_1;E[\mathit{encode}\ttt{(s1,s2)}]$ for certified $\mathit{encode}$, let:
  $$
  \config{\prog_1}{E[\mathit{encode}(\ttt{s1},\ttt{s2})]} \redxs
  \config{\prog_1;\prog_2}{E[v]}
  $$
  If we have the following preconditions, where $\iov(\prog_1) = S' \cup V$:
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$
  \item $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$
  \end{enumerate}
  then we have as a postconditions, where $\iov(\prog_1;\prog_2) = S' \cup V'$: 
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{2\}}}{V'_{\{1\}}}$
  \item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{1\}}}{V'_{\{2\}}}$
  \end{enumerate}
\end{lemma}

\begin{proof}
  By assumptions of well-typedness we have:
  $$
  \begin{array}{rcl}
    v &=&  \ttt{\{shares1 = }v_1\ttt{; shares2 = }v_2\ttt{\}}\\
    v_1 &=& \ttt{\{ c1 = v[1,s1out]; c2 = v[2,s1out] \}}\\
    v_2 &=& \ttt{\{ c1 = v[1,s2out]; c2 = v[2,s2out] \}}
  \end{array}
  $$
  And by Definition \ref{definition-gmwencode-certification} we have:
  \begin{mathpar}
    \sep{\progtt(\prog)}{\ttt{\{s[1,s1]\}}}{\ttt{\{v[2,s1out],v[2,s2out]\}}}

    \sep{\progtt(\prog)}{\ttt{\{s[2,s2]\}}}{\ttt{\{v[1,s1out],v[1,s2out]\}}}
  \end{mathpar}
  Since we assume that secrets are in uniform and independent marginal
  distributions a priori, and $\vars(\prog_1) \cap \vars(\prog_2) =
  \varnothing$ by condition (iii) of Definition 
  \ref{definition-gmwencode-certification} and assumptions of well-typedness, 
  the result follows by preconditions and Lemma \ref{lemma-separation}.
\end{proof}

\begin{theorem}
  \label{theorem-gmw-NIMO}
  If $\eassign{\outv}{\ttt{decode}(e)}$ is a well-typed GMW circuit
  definition using certified components and
  $\config{\varnothing}{\eassign{\outv}{\ttt{decode}(e)}} \redxs
  \config{\prog}{\varnothing}$, then $\NIMO(\prog)$.
\end{theorem}
\begin{proof}
  Let $\iov(\prog) = S \cup (V \cup \outv)$. By definition of $\ttt{decode}$
  and assumptions of well-typedness we have for some $\ttt{g}$:
  $$
  \config{\varnothing}{\eassign{\outv}{\ttt{decode}(e)}} \redxs
  \config{\prog;\eassign{\outv}{\ttt{v[1,gout] xor v[2,gout]}}}{\varnothing}
  $$
  where by Lemmas \ref{lemma-gmw-encode} and \ref{lemma-gmw-preservation} we
  have $\sep{\progd(\prog)}{S_{\{1\}}}{V_{\{2\}}}$ and
  $\sep{\progd(\prog)}{S_{\{2\}}}{V_{\{1\}}}$. 
\end{proof}
